Background for transient faults in asynchronous circuits

### 2019-09-03 draft_together3.tex
before remove SM consistent

\documentclass[12pt]{report}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{suthesis-2e}
\usepackage[font=small,labelfont=bf]{caption}

@@

Muller circuits is the approach that D. E. Muller originally proposed based on this delay model \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synethsis methods begin with some higher-level specification such as signal transition graph or petri nets and translate into a state graph.  State graphs describes the progression of changes of a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allows the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed independent asynchronous circuit.   
There are other types of delay models and asynchronous circuit synthesis methods such as Quasi Delay Insensitive circuits and the relevant types for this work are described in more detail in the literature review.  \\  %add refs to gen C etc

@@

\subsection{Speed Independent Circuits}
As mentioned in the introduction, the unbounded gate delay model assumes an arbitrary delay on all the gates and ignores any delay on wires.  This delay model is the basis for Muller circuits \cite{Muller_59}.  State graphs describe how we would like the circuit to function.  
Muller circuits are then synthesized from the state graphs.
\begin{definition}A state graph consists of $\langle W, S, T\rangle$ where $W$ is the set of finite signals $\{w_1 .. w_n\}$.  $S$ is the set of functions $[W \to \{0,1\}]$ %<-this might not be right math... and s(k) denotes the (binary) value of the state at signal $v_k$.  
which maps each variable to a boolean.  An element of $S$ is called a {\em state}. $T$ is the transition relation of the state graph, $T \subseteq S \times S$.  \end{definition}

@@

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph have the following properties, it is possible to synthesize a hazard free and speed independent asynchronous circuit following the state graph.  Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire.  We next look at how such a gate logic can be made.\\

@@

The two specific Muller circuit implementations that we consider are the generalized C-element and standard C-implementations.  We assume that the circuits we want to make fault tolerant follow these implementations.    
For each wire in the state graph there is a gate logic that controls the value of the wire.  We look at two conditions, set, the condition for the wire to transition to a high value; and reset, the condition for the wire to transition to a low value.  The gate logic then implements the set and reset conditions. The details of how this is implemented is that we use the state graph to generate the set and reset conditions. For each set and reset there is an ON-set, OFF-set and Don't Care set.  ON-set are all states in the state graph where the wire is excited and a transition will occur.  That is for the set condition, it must evaluate to TRUE for these states so that the wire can transition to high.  OFF-set includes all states where the set condition must evaluate to FALSE which is the states where the wire is excited but to the opposite value, which is low for set, as well as all states in the state graph where the wire value is at the opposite value, again low for set.  The rest of the states are in the Don't Care set which includes states that are not included in the state diagram.  It does not matter what the set condition evaluates to in these states as they do not disturb the function of the circuit. The same procedure applies as well in finding the ON-set, OFF-set and Don't Care set for reset except substituting in reset for set and flipping the value of the wire where needed. The set and reset conditions then must include all states in the ON-set and not include all states in the OFF-set so that the wire transitions according to the underlying state graph.  The set and reset can be implemented separately with logic and the gate logic for the wire includes logic for both set and reset and also the combination of them into the output wire.   \\   %Beerel and Myers.... also expand?

@@

For the standard C-implementation, we allow for more variation in delay.  For the gate logic of each variable in the state graph, we look at the set and reset logic as composition of various gates (such as AND, OR, NOT).  Each of these gates can have their own delay that is arbitrary in length that is instantized as an intermediate wire.  Thus the logic must be synthesized more carefully to eliminate hazards that can occur in these intermediate wires.  In particular the set and reset must be activated at states only when the wire is appropriately excited.  As a result, the standard C-implementation has a number of set logic functions combined by an OR gate and fed into one input of the C-element.  There will also be a number of reset logic functions combined by an OR gate and passed through an inverter and fed into another input of the C-element.  This implementation uses the standard C-element so that if the set and reset inputs are mismatched, the output is held.  In general the synthesis can be changed so that there is a hazard-free implementation using only two-input gates.

@@

We define a {\em trace} to describe a path the circuit can take.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states the trace, which is formally defined as follows:
\begin{definition}A {\em trace} $\sigma=(s_1, s_2, ...)$ of a circuit is a sequence of states of finite or infinite length.  
%Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Every pair of consecutive states in a trace are in a transition relation $T(s_i,s_{i+1})$, where state $s_i$ represents the $i^{th}$ term of the sequence.  .  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
\end{definition} 
We also use $\sigma_n$ as shorthand to denote the sequence of the first $n$ states of $\sigma$.
Taking different paths form different traces so a circuit has a set of traces.  % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
The set of traces that can occur depends on the {\em initial} state of the wires in the circuit and the primary inputs.  When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
 The circuit may be waiting for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
\begin{definition}
A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
\end{definition}
In other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces can describe a circuit completely, in addition we may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state is {\em reachable} if it is in some trace, of a set of traces.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces.  
\end{definition}
Thus, unreachable states are states that are not in any of the traces, of a set of traces.  Since all functionality of a circuit can be described by a set of traces, then the circuit will never enter such a state, hence termed unreachable.

@@

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption{a) original circuit b) after the fault tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

Other than the basic terminology for asynchronous circuits, we also want terminology to compare two circuits based on their function.  
We first define a {\em trace}, to describe a path the circuit can take as it follows the state graph.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states and transitions the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of finite or infinite length and $s_i$ is the $i^{th}$ term of the sequence.  %Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Also every pair of consecutive states in a trace are transitions.  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
We also use the $\sigma_n$ as shorthand to denote the sequence of the first n states of $\sigma$\end{definition} 
Thus a circuit has a set of traces.  A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.\\%new 4/30

Note that we allow the circuit to have external wire inputs and stopping states.  
\begin{definition}A {\em stopping state} is a state where there are no excitations in the wires in the circuit and the circuit is waiting for the external inputs.\end{definition} 
A trace of finite length ends in a stopping state where the external inputs never arrive and the circuit is stopped.  A related concept is a deadlocked state which is a state where the circuit prematurely stops and defined as follows:
\begin{definition}A {\em deadlocked state} is a state where there are no excitations in the wires in the circuit and is not a stopping state.\end{definition}
A deadlocked state does not appear in a well designed circuit, but may appear after a circuit transformation (i.e., the new circuit stops at states where the original circuit do not), or when there are faults.  Thus we want to design a circuit transformation where there are no deadlocked states.\\ %end new
%Note that our definition of trace also includes the case where there is only one transition from each state of a state graph, which results in a linear state graph.

@@

The original circuit is semi-modular by design.  But the design only covers cases where the wires are in a state in the state graph.  That is, hazards may occur if the circuit goes into an unreachable state of the state graph and the circuit is no longer semi-modular and furthermore becomes unpredictable.  This does not happen in the fault-free case where the circuit is initialized to a state in the state graph and the wires inside the logic blocks are properly initialized.  However when faults do occur, the circuit may become unpredictable which is a reason why we need to transform the circuit to protect against faults.  An example of what might happen when the circuit is not in a semi-modular state is shown in figure \ref{fig:sm_counter2}.  The logic circuit for wire $w$ is taken out of a semi-modular circuit.  However, if the state of the $w,x,y,z$ wires are $(0,1,0,1)$ and the internal wires after the AND and OR gate are both initialized to 1, this is not semi-modularity consistent.  The output wire $w$ is excited to 1 while the inputs from $x,y,z$ will cause $w$ to transition to 0.  Depending on the timing of the gates, $w$ may stay at 0 or experience a glitch to 1 then back to 0.  The extra layers of gates in the logic block can be thought of as extra delays that can store and propagate the wrong values.  More detailed discussion on how to initialize and keep the circuit semi-modular is explored in the subsequent section.\\

@@

We want to show that a half circuit from the fault tolerant circuit has correct function.  To do this, we define a correspondence between states of the original circuit and states of the fault tolerant circuit.  As a result of the circuit transformation, a state of the fault tolerant circuit has bits for 8 times as many wires as a state in the original circuit.  Half of those bits are obviously from the A half and half are from the B half of the circuit.  The state graph for a fault tolerant circuit is defined on this combined state.  We define the half-circuit mapping $h^A: s^{\mathit{full}} \to s^A$ where $s^{\mathit{full}}$ is the combined state of a fault tolerant circuit and $s^A$ consists of wires $w^A$ that correspond to the original wires $w$ in the original duplication procedure.  The half-circuit mapping $h^B$ is similarly defined to map $s^{\mathit{full}}$ to $s^B$.  The added wires such as $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ from the duplication procedure are hidden by $h^A$ or $h^B$.  We extend $h^A$ to traces by using the following two steps: first, $h^A$ is applied individually to each state in a trace $\sigma$, and likewise for $h^B$.  %not sure if I want the mapping to be from variable to variable or specific state to state
%For a trace in the fault tolerant circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). 
The resulting trace may have successive repeated states.  So next we delete all but one of the neighboring repeats to obtain a valid trace on the shortened list of wires.  The result should be a trace from the original circuit.  
Using this mapping we will show that the original circuit is equivalent to half of the fault tolerant circuit in both the absence and presence of a single transient fault.
%do I need to define all the new transitions in duplicated circuit?  maybe
\\

%[Why is this the best we can do?]
We compare each half circuit because it is not easy to synchronize and compare the signals between both halves, especially if we reuse the logic blocks from the original circuit.  A simple idea we tried was to insert another C-element before the input is fed into the logic, to synchronize the wires on both halves.  The problem with that is in the original circuit, a wire value change propagates to all inputs at the same time.  Adding another layer of C-elements creates delays at the input to logic blocks and the isochronic fork assumption is violated.  Thus the resulting circuit will not function according to the specified state graph.  (an example can be provided but not sure if needed?)\\  
\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{counterexhalf}
  \caption{State graph example to show how b can transition before a or c finishes in both halves in the fault tolerant circuit}
  \label{fig:counterexhalf}
\end{figure}

Next we examine the traces of each half circuit.  Figure \ref{fig:counterexhalf} shows a subsection of a semi-modular state graph where each state is given a nominal name (state 0, state 1, ... etc) and each transition is labeled with the transitioning wire.  Suppose we apply our fault tolerant transformation to such a circuit based on this state graph.  Both halves in the fault tolerant circuit will transition based on this state graph, with the added caveat that for a wire to transition at some state, it must be excited in both halves.  Starting in {\em state 0}, it is possible for the A-half to transition to {\em state 1} ($a^A$ transitions) and for the B-half to transition to {\em state 2} ($c^A$ transitions).  We assume that $a^B$ and $c^A$ do not complete their transitions yet.  Immediately, in the A-half, $b^A$ is excited and in the B-half, $b^B$ is excited.  Then, suppose the transition on wire $b^A$ and $b^B$ occurs.  That is, the A-half transitions to {\em state 3} and the B-half transitions to {\em state 5}.  And finally, the excited $a^B$ and $c^A$ transitions so that both halves are in {\em state 6}.  The trace so far for the A-half is (state 0, state 1, state 3, state 6) while for the B-half it is (state 0, state 2, state 5, state 6).  Thus the two halves need not have the same trace.  One criticism may be that the current fault tolerant transformation synchronizes at the outputs of the logic blocks.  If we want to reuse the same logic blocks but synchronize the signals earlier, such as replacing the final C-element in a standard C implementation with a 4C-element, the issue of different traces in the two halves will still occur.  We have already shown earlier that synchronizing the input signals while keeping the original logic block can also result in error.  This points to looking at properties of each half circuit as the right thing to do.\\

@@

\section{Normal operation analysis}
Now we will discuss the proper wire initializations before going into the Lemmas.  In normal operation, the fault tolerant circuit must have wire values consistent with semi-modular behavior in the original circuit.  
This includes the internal wires which exist in the logic blocks in the original circuit and through duplication also exist in the fault tolerant circuit.  The actual values of these wires do not matter if the inputs to the logic blocks follow the semi-modularity property, and if the internal wires are initialized properly.  This arises from the original synthesis method to synthesize the circuit from an underlying state graph.  
Additionally, it does not suffice for the inputs to the logic blocks to be from a valid trace of the state graph; the timing matters as well.  The semi-modularity property must be preserved:  if the logic block output is excited then it must remain excited until a transition on its output occurs.  Since certain inputs which are valid states and arise from valid transitions may cause the logic block to become unexcited before a transition occurs and thus break the semi-modularity property, we require this more stringent criteria.  %maybe needs an example
Thus to satisfy semi-modularity, the inputs wait at certain input states until the output transitions.  Then the internal wires of the logic block are termed {\em semi-modularity consistent} if the following properties are true:  if the inputs follow semi-modularity then there are no glitches at the output; and once the output is excited it remains excited until the output transitions. \\%newly added

>>>>

\documentclass[12pt]{report}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{suthesis-2e}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{comment}

@@

Muller circuits is the approach that D. E. Muller originally proposed based on this delay model \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synethsis methods begin with some higher-level specification such as signal transition graph or petri nets and translate into a state graph.  State graphs describes the progression of changes of a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allows the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed independent asynchronous circuit.  We assume that the previous steps are hidden, and are presented with a semi-modular circuit to make fault tolerant.     
There are other types of delay models and asynchronous circuit synthesis methods such as Quasi Delay Insensitive circuits and the relevant types for this work are described in more detail in the literature review.  \\  %add refs to gen C etc

@@

\subsection{Speed Independent Circuits}
As mentioned in the introduction, circuits are composed of gates and wires.  The unbounded gate delay model assumes an arbitrary delay on all the gates and no delay on wires.  This delay model is the basis for Muller circuits \cite{Muller_59}.  Circuits that operate correctly regardless of the delay are called Speed Independent Circuits.  The correct operation is specified when designing the circuit, for example, given some input values the circuit will return specified output values.  
There are many ways to represent specifications such as petri nets and state graphs, and there are established methods to build asynchronous circuits to satisfy the specifications.  We won't go into these representations and methods in detail, other than the aspects that are relevant to this work.  We consider circuits which has wires that take on Boolean values of either `1' or `0'.\\
%Wires in the circuits take on Boolean values of either `1' or `0'.  A state is defined by assigning a Boolean value to each wire of the circuit.  States are unique in that no two states share the same wire values.  There will be a formal definition of wires and states in the next chapter.

\begin{comment}
State graphs describe how we would like the circuit to function.  
Muller circuits are then synthesized from the state graphs.
\begin{definition}A state graph consists of $\langle W, S, T\rangle$ where $W$ is the set of finite signals $\{w_1 .. w_n\}$.  $S$ is the set of functions $[W \to \{0,1\}]$ %<-this might not be right math... and s(k) denotes the (binary) value of the state at signal $v_k$.  
which maps each variable to a boolean.  An element of $S$ is called a {\em state}. $T$ is the transition relation of the state graph, $T \subseteq S \times S$.  \end{definition}

@@

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph have the following properties, it is possible to synthesize a hazard free and speed independent asynchronous circuit following the state graph.  Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire.  We next look at how such a gate logic can be made.\\ 
\end{comment}

@@

The two specific Muller circuit implementations that we consider are the generalized C-element and standard C-implementations.  We assume that the circuits we want to make fault tolerant are synthesized according to these two implementations.    
Each wire in the circuit is the output wire of a gate logic.  The gate logic has a set of input wires and the output wire changes value according to the values of the inputs.  We look at two conditions, set, the condition for the output wire to transition to a high value; and reset, the condition for the output wire to transition to a low value.  The gate logic then implements the set and reset conditions. 
We skim the details in finding the set and reset conditions from the circuit specifications. For each condition set and reset, there is an ON-set, OFF-set and Don't Care set.  ON-set are all input values where the wire is excited and a transition will occur.  The OFF-set includes all input values where the wire cannot be excited to some value per the circuit specifications.  Input values that are neither in the ON-set or OFF-set are in the Don't Care set.  So for the set condition, the ON-set are all input values where the wire must transition to high.  The OFF-Set includes all input values where the wire is excited instead to low, as well as input values for which the output wire must remain low.  The input values of the Don't Care set assign the output wire to be either high or low.  It does not matter which value is assigned since it does not disturb the function of the circuit and only helps for circuit optimization. \\

The same procedure applies as well in finding the ON-set, OFF-set and Don't Care set for reset except substituting in reset for set and flipping the value of the wire from high to low and vice versa. The set and reset conditions constructed this way then follows the desired circuit specification.  The set and reset conditions are implemented using gates and are combined into the output wire through a C-element.  This entire block of gates and C-element for the output wire is called the gate logic. \\   %Beerel and Myers.... also expand?

@@

For the standard C-implementation, we allow for more variation in delay.  For each wire in the circuit with a corresponding gate logic, we look at the set and reset logic as composition of various gates (such as AND, OR, NOT).  Each of these gates can have their own delay that is arbitrary in length that is instantized as an intermediate wire.  Thus the logic must be synthesized more carefully to eliminate hazards that can occur in these intermediate wires.  In particular the set and reset must be activated at states only when the wire is appropriately excited.  As a result, the standard C-implementation has a number of set logic functions combined by an OR gate and fed into one input of the C-element.  There will also be a number of reset logic functions combined by an OR gate and passed through an inverter and fed into another input of the C-element.  This implementation uses the standard C-element so that if the set and reset inputs are mismatched, the output is held.  In general the synthesis can be changed so that there is a hazard-free implementation using only two-input gates.

@@

We define a {\em trace} to describe a path the circuit can take.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states the trace, which is formally defined as follows:
\begin{definition}A {\em trace} $\sigma=(s_1, s_2, ...)$ of a circuit is a sequence of states of finite or infinite length.  
%Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Every pair of consecutive states in a trace are in a transition relation $T(s_i,s_{i+1})$, where state $s_i$ represents the $i^{th}$ term of the sequence.  .  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
\end{definition} 
We also use $\sigma_n$ as shorthand to denote the sequence of the first $n$ states of $\sigma$.
When there multiple transitions from a state, multiple traces exist depending on which wire transitions.  Thus we want to define a set of traces for all the traces that form.  In addition the initial state is important in naming a trace and by extension for a set of traces.  The {\em initial state} specifies the state of the wires in the circuit as well as the primary inputs.  Thus a set of traces is defined as follows:
\begin{definition}
A {\em set of traces} of a circuit with initial state $s_1$ is the set of all traces of the circuit starting in state $s_1$. 
\end{definition}

 % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
 The circuit may be waiting for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
\begin{definition}
A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
\end{definition}
In other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces can describe a circuit completely, in addition we may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state $s$ is {\em reachable} from $s_1$ if $s$ appears in some trace that begins with $s_1$.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces. 
%still kind of weird oh well 
\end{definition}
Thus, unreachable states are states that are not in any of the traces with initial state $s_1$.  Since all functionality of a circuit can be described by a set of traces starting in a specified state, then the circuit will never enter those unreachable states.

@@

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all of its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption{a) original circuit b) after the fault tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

\begin{comment}
Other than the basic terminology for asynchronous circuits, we also want terminology to compare two circuits based on their function.  
We first define a {\em trace}, to describe a path the circuit can take as it follows the state graph.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states and transitions the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of finite or infinite length and $s_i$ is the $i^{th}$ term of the sequence.  %Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Also every pair of consecutive states in a trace are transitions.  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
We also use the $\sigma_n$ as shorthand to denote the sequence of the first n states of $\sigma$\end{definition} 
Thus a circuit has a set of traces.  A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.\\%new 4/30

Note that we allow the circuit to have external wire inputs and stopping states.  
\begin{definition}A {\em stopping state} is a state where there are no excitations in the wires in the circuit and the circuit is waiting for the external inputs.\end{definition}

In the previous chapter, we have introduced the trace of a circuit, stopping state and complete trace.

A trace of finite length ends in a stopping state where the external inputs never arrive and the circuit is stopped.  A related concept is a deadlocked state which is a state where the circuit prematurely stops and defined as follows:
\begin{definition}A {\em deadlocked state} is a state where there are no excitations in the wires in the circuit and is not a stopping state.\end{definition} %but stopping state= no excitations, so either does not introduce new stopping states after the transformation or what....
A deadlocked state does not appear in a well designed circuit, but may appear after a circuit transformation (i.e., the new circuit stops at states where the original circuit do not), or when there are faults.  Thus we want to design a circuit transformation where there are no deadlocked states.\\ %end new

\end{comment}
%Note that our definition of trace also includes the case where there is only one transition from each state of a state graph, which results in a linear state graph.

@@

The original circuit is semi-modular by design.  But the design only covers traces when the circuit starts in a specified initial state.  That is, hazards may occur if the circuit goes into an unreachable state of the traces and the circuit is no longer semi-modular and furthermore becomes unpredictable.  This does not happen in the fault-free case and the circuit starts in a valid initial state.  However when faults do occur, the circuit may become unpredictable, which is a reason why we need to transform the circuit to protect against faults.  An example of what might happen when the circuit is not in a semi-modular state is shown in figure \ref{fig:sm_counter2}.  The logic circuit for wire $w$ is taken out of a semi-modular circuit.  However, if the state of the $w,x,y,z$ wires are $(0,1,0,1)$ and the internal wires after the AND and OR gate are both initialized to 1, this is not semi-modularity consistent.  The output wire $w$ is excited to 1 while the inputs from $x,y,z$ will cause $w$ to transition to 0.  Depending on the timing of the gates, $w$ may stay at 0 or experience a glitch to 1 then back to 0.  The extra layers of gates in the logic block can be thought of as extra delays that can store and propagate the wrong values.  More detailed discussion on how to initialize and keep the circuit semi-modular is explored in the subsequent section.\\

@@

We want to show that a half circuit from the fault tolerant circuit has correct function.  To do this, we define a correspondence between states of the original circuit and states of the fault tolerant circuit.  As a result of the circuit transformation, a state of the fault tolerant circuit has bits for 8 times as many wires as a state in the original circuit.  Half of those bits are obviously from the A half and half are from the B half of the circuit.  The trace of a fault tolerant circuit is defined on this combined set of wires.  We define the half-circuit mapping $h^A: s^{\mathit{full}} \to s^A$ where $s^{\mathit{full}}$ is the combined state of a fault tolerant circuit and $s^A$ consists of wires $w^A$ that correspond to the original wires $w$ in the original duplication procedure.  The half-circuit mapping $h^B$ is similarly defined to map $s^{\mathit{full}}$ to $s^B$.  The added wires such as $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ from the duplication procedure are hidden by $h^A$ or $h^B$.  We extend $h^A$ to traces by using the following two steps: first, $h^A$ is applied individually to each state in a trace $\sigma$, and likewise for $h^B$.  %not sure if I want the mapping to be from variable to variable or specific state to state
%For a trace in the fault tolerant circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). 
The resulting trace may have successive repeated states.  So next we delete all but one of the neighboring repeats to obtain a valid trace on the shortened list of wires.  The result should be a trace from the original circuit.  
Using this mapping we will show that the original circuit is equivalent to half of the fault tolerant circuit in both the absence and presence of a single transient fault.
%do I need to define all the new transitions in duplicated circuit?  maybe
\\

%[Why is this the best we can do?]
We compare each half circuit because it is not easy to synchronize and compare the signals between both halves, especially if we reuse the logic blocks from the original circuit.  A simple idea we tried was to insert another C-element before the input is fed into the logic, to synchronize the wires on both halves.  The problem with that is in the original circuit, a wire value change propagates to all inputs at the same time.  Adding another layer of C-elements creates delays at the input to logic blocks and the isochronic fork assumption is violated.  Thus the resulting circuit will not function according to the desired specifications.  (an example can be provided but not sure if needed?)\\  
\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{counterexhalf}
  \caption{State graph example to show how b can transition before a or c finishes in both halves in the fault tolerant circuit}
  \label{fig:counterexhalf}
\end{figure}

Next we examine the traces of each half circuit.  Assume that for the original circuit starting in state 0, both (state 0, state 1, state 3, state 6) and (state 0, state 2, state 5, state 6) are prefixes of valid traces.  Then both halves in the fault tolerant circuit will transition based on the traces of the original circuit, with the added caveat that for a wire to transition at some state, it must be excited in both halves.  Since only one wire transitions at a time, it is possible for say $a^A$ to transition first and $a^B$ to transition later.  Starting in {\em state 0}, it is possible for the A-half to transition to {\em state 1} ($a^A$ transitions) and for the B-half to transition to {\em state 2} ($c^B$ transitions).  We assume that $a^B$ and $c^A$ do not complete their transitions yet.  Immediately in the A-half, $b^A$ is excited, and in the B-half, $b^B$ is excited.  Since $b$ is excited in both halves, wires $b^A$ and $b^B$ can transition.  That is, the A-half transitions to {\em state 3} and the B-half transitions to {\em state 5}.  And finally, the excited $a^B$ and $c^A$ transitions so that both halves are in {\em state 6}.  The trace so far for the A-half is (state 0, state 1, state 3, state 6) while for the B-half it is (state 0, state 2, state 5, state 6).  Thus the two halves need not have the same trace.  One criticism may be that the current fault tolerant transformation synchronizes at the outputs of the logic blocks.  If we want to reuse the same logic blocks but synchronize the signals earlier, such as replacing the final C-element in a standard C implementation with a 4C-element, the issue of different traces in the two halves will still occur.  We have already shown earlier that synchronizing the input signals while keeping the original logic block can also result in error.  This points to looking at properties of each half circuit as the right thing to do.\\

@@

\section{Normal operation analysis}
Now we will discuss the proper wire initializations before going into the Lemmas.  Defining a set of traces of a circuit includes defining the initial state.  For the original circuit, if the circuit starts from this initial state, then the synthesis method ensures semi-modularity in all traces.  However, the proposed fault tolerant transformation adds logic, and the set of traces and initial state may be different from those of the original circuit, even after applying the half circuit mappings.  Thus we need to ensure that after the transformation, the circuit behaves in a similar way to the original circuit.  That is for an initial state of the fault tolerant circuit, which is related to the initial state of the original circuit, the resulting traces will behave similarly and have properties such as semi-modularity and the same stopping states as the original circuit. \\

%the half circuits are still semi-modular, has the same transitions as the original circuit and does not stop in non-stopping states of the original circuit. 
%This initial state consistent with semi-modularity also includes the initial values of internal wires.

We show how to define the initial state for the fault tolerant circuit if given the initial state for the original circuit.  The values for the internal wires must also be set.  These internal wires are initialized to a semi-modularity consistent value.  Given this, we pick the following convenient initializations for the fault tolerant circuit.  The following procedure covers how to initialize both the wires from the duplication as well as the extra $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ wires that are introduced.
%For all initial states we assume that we know the values of every wire in the circuit, intermediate and visible wires.  The visible wire values are the output wires in the circuit $w$, and the wires $c1$, $c2$ and $c3$'s. The intermediate wires are all other wires, which consists of all the wires within the logic blocks for each output wire.  Given these we will show how to define a corresponding initial state in the duplicated circuit.  %should I delete intermediate?
\begin{definition} A {\em valid initial state} for the A half of the fault tolerant circuit is constructed from the initial state of the original circuit.  The wire values in the logic block and $w$ in the original circuit are copied to the corresponding wires in the A half.  The additional wires from the C-element chain are also assigned the same value as $w$.  That is, first assign the value of $w$ to $w^A$, then assign the value of $w$ to $c1^A$ $c2^A$ $c3^A$ as well. The same construction applies for the B half circuit as well.
\end{definition}
%\textbf{
This definition is stronger than it needs to be but is convenient.\\

\begin{comment}
%dynamic properties of SM?  I don't think it's needed right now
The input wire values must also satisfy dynamic properties.  Static properties such as the input wire values must be from valid states of the trace is not enough.  The order and the wire transitions also matter.   
The semi-modularity property must be preserved: if the logic block output is excited then it must remain excited until a transition on its output occurs.  Since certain inputs which are valid states and arise from valid transitions may cause the logic block to become unexcited before a transition occurs and thus break the semi-modularity property, we require this more stringent criteria.
